---
title: "analysis"
output: html_document
---
# load packages
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, modelsummary, fixest)
```


# Question 1: Uncompensated Care Over Time
```{r}
data %>% 
  group_by(year) %>% 
  filter(!is.na(tot_uncomp_care_charges)) %>%
  filter(year %in% c(2010:2018)) %>% 
  summarize(uncomp_care = mean(tot_uncomp_care_charges)/1000000) %>% 
  ggplot(aes(x = year, y = uncomp_care)) +
  geom_line() + 
  labs(title = "Average Uncompensated Care Charges Over Time",
       x = "Year",
       y = "Average Uncompensated Care Charges (in millions)")
```

# Question 2: Uncompensated Care in Never Expanded States vs. 2014 Adopters
```{r}
reg_dat_2014 %>% 
  group_by(year, expanded) %>% 
  filter(year %in% c(2010:2018)) %>%
  summarize(uncomp_care = mean(tot_uncomp_care_charges)/1000000) %>% 
  ggplot(aes(x = year, y = uncomp_care, color = as.factor(expanded))) + 
  geom_line() +
  labs(title = "Average Uncompensated Care Charges Over Time by Medicaid Expansion Status",
       x = "Year",
       y = "Average Uncompensated Care Charges (in millions)",
       color = "Medicaid Expansion Status") +
  scale_color_discrete(
    labels = c(`FALSE` = "Never expanded", `TRUE` = "Expanded in 2014")
  )
```
# Question 3:  2x2 DD table 
```{r}
reg_dat_2014 %>%  
  filter(year %in% c(2012, 2015)) %>% 
  group_by(year, expanded) %>%
  mutate(expanded = if_else(expanded,
    "2014 expanders",
    "Never expanded"
  )) %>% 
  summarize(uncomp_care = mean(tot_uncomp_care_charges)/1000000) %>% 
  pivot_wider(names_from = year, values_from = uncomp_care) %>% 
  rename(Pre = `2012`, Post = `2015`, Group = expanded)
```
# Question 5: DD for 2014 adopters vs. no adoption
```{r}
dd_2014 <- lm(tot_uncomp_care_charges/1000000 ~ post + expanded + treat, data = reg_dat_2014)
modelsummary(dd_2014, statistic = "p.value", stars = TRUE)
```

# Question 6: TWFE for 2014 adopters vs. no adoption
```{r}
twfe_2014 <- feols(tot_uncomp_care_charges/1000000 ~ treat | provider_number + year, 
                   data = reg_dat_2014,
                   cluster = ~ state)
modelsummary(twfe_2014, statistic = "p.value", stars = TRUE)
```

# Question 7: TWFE for all adopters vs. no adoption
```{r}
# make dataset for twfe
twfe_dat_all <- data %>% 
  filter(!is.na(tot_uncomp_care_charges)) %>%
  mutate(
    year_adopted = year(date_adopted),
    post = (year >= year_adopted),
    treat = post * expanded
  )

# run model
twfe_all <- feols(tot_uncomp_care_charges/1000000 ~ treat | provider_number + year, 
                   data = twfe_dat_all,
                   cluster = ~ state)
modelsummary(twfe_all, statistic = "p.value", stars = TRUE)
```

# Question 8: Event Study Graph for 2014 adopters vs. no adoption
```{r}
event_study_2014 <- feols(tot_uncomp_care_charges/1000000 ~ i(year, expanded, ref = 2013) | provider_number + year, 
                   data = reg_dat_2014,
                   cluster = ~ state)
iplot(event_study_2014,
      xlab = "Year",
      main = "Event Study: Uncompensated Care for 2014 Adopters vs. No Adoption") +
  geom_vline(xintercept = 2014, linetype = "dashed", color = "red") 
```
# Question 9: Event Study Graph for all adopters vs. no adoption
```{r}
event_dat_all <- data %>% 
  filter(!is.na(expanded)) %>%
  mutate(
    time_to_treat = if_else(expanded == FALSE, 0, year - year(date_adopted)),
    time_to_treat = if_else(time_to_treat < -3, -3, time_to_treat)
  )

event_study_all <- feols(tot_uncomp_care_charges/1000000 ~ i(time_to_treat, expanded, ref = -1) | provider_number + year,
                       cluster = ~ state,
                       data = event_dat_all)
iplot(event_study_all,
      xlab = "Years Relative to Adoption",
      main = "Event Study: Uncompensated Care for All Adopters vs. No Adoption") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "red")
```

